A full approximation storage multilevel method \\ for nonlinear variational inequalities

### 2022-12-23 paper/fascd.tex
rename files: mcd2->fascd
edits through start section 8

\begin{lemma}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable, and let $u\in\cK$.  For $0<\alpha\le 1$ the output $w \in \cK$ from \pr{cd-mult} satisfies
\begin{equation}
F(w) - \ip{\ell}{w} \le F(u) - \ip{\ell}{u}.  \label{eq:monotone}
\end{equation}
If $0 < \alpha \le 1/m$ then \pr{cd-add} likewise satisfies \eqref{eq:monotone}.
\end{lemma}

@@

The \fascd V-cycle algorithm in this section uses Lemmas \ref{lem:downwardadmissibility} and \ref{lem:upwardadmissibility}, that is, it smooths in $\mathcal{D}^j$ during the downward stages, and in $\mathcal{U}^j$ for the upward stages, respectively.  The algorithm extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear FAS approach \cite{BrandtLivne2011}.  We coarsen the solution iterate itself, using injection, in constrast to classical linear multigrid wherein only the residual is coarsened.

The finest-level iterate $w^J \in \mathcal{K}^J$ will be corrected by perturbations from each coarser subspace $\mathcal{V}^j$,\footnote{By nesting \eqref{eq:fe:nestedspaces} these corrections also live in the fine-level subspace $\mathcal{V}^J$.  Later we will use prolongation $P$ to indicate where a change in computer representation would be needed.} but the corrected iterate needs to be admissible.  In detail, suppose $y^J \in \mathcal{D}^J, y^{J-1} \in \mathcal{D}^{J-1}, \dots, y^{j+1} \in \mathcal{D}^{j+1}$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:nmcdvcycle}).  By \eqref{eq:fe:downwardsumcd} the corrected iterate at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.  Smoothing then happens at the next-coarser level $j$, i.e.~we incompletely-solve VI problem \eqref{eq:fe:downvi} below.  At the bottom the coarsest-level VI is then (accurately) solved, to generate the full downslash correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.  Starting upward, however, we can use $y^1+ z^0 \in \mathcal{U}^1$ as an initial iterate on level $j=1$, and then smooth by incompletely-solving VI problem \eqref{eq:fe:upvi} (below) in $\mathcal{U}^1$, yielding admissible total correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsumcd}.  Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$ from initial iterate $y^j+z^{j-1}$, eventually generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction, yielding the new iterate $w^J + z^J \in \mathcal{K}^J$.

\begin{figure}[ht]
\begin{center}
\input{tikz/fascdvcycle.tex}
\end{center}
\caption{A \fascd V-cycle computes downward corrections $y_j \in \mathcal{D}^j$ and upward corrections $z_j\in\mathcal{U}^j$.  The latter live in larger constraint subsets.}
\label{fig:fascdvcycle}
\end{figure}

A key idea, implicit in the above paragraph, is that the up-smoothing stages solve less-restricted VI problems than down-smoothing stages.  Intuitively-speaking, downward one does not know what the coarse corrections will be, so they must be in such small sets ($\mathcal{D}^j$) that their sum is admissible once we return to the current level.  However, going upward smoothing can occur in the larger admissible-sum sets ($\mathcal{U}^j$; see \eqref{eq:fe:constraintdecomp}) because there are no forthcoming coarser corrections which could violate admissibility.

@@

\bibliography{mcd2}
\bibliographystyle{siam}

\appendix
\section{Reductions of NMCD to familiar multilevel algorithms}

>>>>

\begin{lemma}  Suppose $f=F'$ for $F:\cK\to\RR$ convex and differentiable, and let $u\in\cK$.  For $0<\alpha\le 1$ the output $w \in \cK$ from \pr{cd-mult} satisfies
\begin{equation}
F(w) - \ip{\ell}{w} \le F(u) - \ip{\ell}{u}.  \label{eq:objectivemonotone}
\end{equation}
If $0 < \alpha \le 1/m$ then \pr{cd-add} likewise satisfies \eqref{eq:objectivemonotone}.
\end{lemma}

@@

The \fascd V-cycle algorithm in this section uses Lemmas \ref{lem:downwardadmissibility} and \ref{lem:upwardadmissibility}, that is, it smooths in $\mathcal{D}^j$ and $\mathcal{U}^j$ during the downward and upward stages, respectively.  \fascd extends the multilevel CD schemes of Tai \cite{Tai2003} and Gr\"aser \& Kornhuber \cite[Algorithm 4.7]{GraeserKornhuber2009} by following a nonlinear FAS approach \cite{BrandtLivne2011}.  Note that we coarsen both the residual and the solution iterate itself, in constrast to classical linear multigrid wherein only the residual is coarsened.  In fact, \fascd reduces to FAS for PDEs when inequality constraints are removed (Appendix \ref{app:reductions}).

The V-cycle generates a correction to a finest-level iterate $w^J \in \mathcal{K}^J$ based on perturbations from each coarser FE subspace $\mathcal{V}^j$.  By nesting \eqref{eq:fe:nestedspaces} these corrections also live in the fine-level subspace $\mathcal{V}^J$, but we will use prolongation $P$ to indicate where the computer representation changes.

However, the corrected iterate needs to be admissible.  In detail, suppose $y^i \in \mathcal{D}^i$ for $i=J,\dots,j-1$ are already-computed corrections during the downward part of the V-cycle (Figure \ref{fig:fascdvcycle}).  By \eqref{eq:fe:downwardsuminclusion} the corrected iterate at this point is admissible, namely $y^J + \dots + y^{j+1} \in \mathcal{U}^J$, equivalently $w^J + y^J + \dots + y^{j+1} \in \mathcal{K}^J$.  Smoothing, an incomplete solution process on a coarser-level VI problem (defined below), then occurs at the next-coarser level, yielding the next correction $y^j \in \mathcal{D}^j$.  At the bottom the coarsest-level VI is then solved, resulting in a downward correction $y^J + \dots + y^1 + z^0 \in \mathcal{U}^J$.  Starting upward, $y^1 + z^0 \in \mathcal{U}^1$ is an initial iterate for smoothing on level $j=1$.  Smoothing in $\mathcal{U}^1$ yields admissible correction $y^J + \dots + y^2 + z^1 \in \mathcal{U}^J$ using \eqref{eq:fe:upwardsuminclusion}.  Continuing in this ``telescoping'' way, up-smoothing in the constraint sets $\mathcal{U}^j$, starting from initial iterate $y^j+z^{j-1}$, finally generates $z^J\in \mathcal{U}^J$ as the full V-cycle correction.  Thus the next admissible iterate is $w^J + z^J \in \mathcal{K}^J$.

\begin{figure}[ht]
\begin{center}
\input{tikz/fascdvcycle.tex}
\end{center}
\caption{The \fascd V-cycle computes downward corrections $y_j \in \mathcal{D}^j$, but the upward corrections $z_j\in\mathcal{U}^j$ are in larger sets.}
\label{fig:fascdvcycle}
\end{figure}

A key idea is that, because $\mathcal{U}^j\supset \mathcal{D}^j$, up-smoothing solves less-restricted VI problems than down-smoothing.  Intuitively-speaking, going downward one does not know what the upcoming coarse corrections will be, so they must be in the sufficiently-small sets $\mathcal{D}^j$ so that their sum is admissible once we return to a level.  However, going upward no further \emph{coarser} corrections, relative to that level, can violate admissibility.

@@

\bibliography{fascd}
\bibliographystyle{siam}

\appendix
\section{Reductions of NMCD to familiar multilevel algorithms} \label{app:reductions}
